Patterns of floral and chromosome evolution in Onagraceae: a fossil-calibrated supermatrix analysis

### 2014-05-14 manuscript.tex
removed elsevier preprint
intro work

@@

\journal{IB200}

@@

Introduce supermatrix analyses....

Onagraceae, the evening primroses, are a well-studied family
of flowering plants with 22 described genera and around 660 species \citep{wagner2007revised}.
The family is found on all continents except Antartica,
and pollen fossils date Onagraceae to the Upper Cretaceous \citep{grimsson}.
Onagraceae is well placed in the order Myrtales,
and previous molecular phylogenetic studies find it most closely related to Lythraceae \citep{conti1997, Sytsma2004}.
Though there have been many genus-level molecular phylogenetic studies in Onagraceae
\citep{Berry2004, Evans2009, Hoggard2004, Xie2009, Baum1994, Wagner2005}, 
the most recent family-level analyses \citep{Levin2003, Levin2004}
used only one or two exemplars from some genera. 
Furthermore, the historical biogeography, divergence times of the major clades, and 
family-level patterns and processes of diversification remain poorly understood.
With the wealth of data available, Onagraceae is an
ideal group to apply a data-mining supermatrix approach
with the goal of constructing a densely sampled family-level phylogeny.

Here, I use a well-resolved supermatrix phylogeny of Onagraceaea to
answer qestions such as:
1) Are the taxonomic groups described in \citet{wagner2007revised} monophyletic?
2) When did the major clades diverge?
3) What is the pattern of chromosome number
evolution?
4) What are the patterns of petal color and petal number evolution, and are their 
evolution correlated?
Finally, I also discuss ways in which these analyses can be extended
to further examine the patterns and processes of
evolution in Onagraceae.

>>>>

% to remove elsearticle 'preprint journal'
\makeatletter
\def\ps@pprintTitle{%
 \let\@oddhead\@empty
 \let\@evenhead\@empty
 \def\@oddfoot{\centerline{\thepage}}%
 \let\@evenfoot\@oddfoot}
\makeatother
%%\journal{IB200}

@@

@@

Biologists are increasingly using massive phylogenetic
datasets to reconstruct ever-growing portions of the Tree of Life 
in pursuit of large-scale evolutionary questions.
A common approach to building large phylogetic trees has been
the supermatrix approach, in which multiple datasets are
combined and analyzed simultaneously \citep{DeQueiroz2007}.
However, assembling and utilizing supermatrices is not trivial
due to difficulties such as determining homology of molecular
sequences, defining operational taxonomic units, 
and managing the amount of missing data \citep{Roure2007, Yan2005}.
phlawd and phylota...
Supermatrix approaches have been successfully used in a number
of large scale angiosperm evolutionary studies including, bling, blang, blot.
In this study I use a supermatrix to build a phylogeny of
the family Onagraceae, and implement a novel method for supermatrix
construction in a Python package.

Onagraceae, the evening primrose family, is a well-studied family
of flowering plants with 22 described genera and around 660 species \citep{wagner2007revised}.
The family is found on all continents except Antartica,
and pollen fossils date Onagraceae to the Upper Cretaceous \citep{grimsson}.
Onagraceae falls within the order Myrtales,
and previous molecular phylogenetic studies find it most closely related to Lythraceae \citep{conti1997, Sytsma2004}.
Though there have been many genus-level molecular phylogenetic studies in Onagraceae
\citep{Berry2004, Evans2009, Hoggard2004, Xie2009, Baum1994, Wagner2005}, 
the most recent family-level analyses \citep{Levin2003, Levin2004}
used only one or two exemplars from some genera. 
Furthermore, the historical biogeography, divergence times of the major clades, and 
family-level patterns and processes of diversification remain poorly understood.
With over 5000 DNA sequences available in GenBank from previous studies, Onagraceae is an
ideal group to apply a data-mining supermatrix approach
with the goal of constructing a densely sampled family-level phylogeny.

Here, I use a well-resolved supermatrix phylogeny of Onagraceae to
test whether the taxonomic groups described in \citet{wagner2007revised} are monophyletic,
and to answer the following questions:
1) When did the major clades diverge?
2) What is the pattern of chromosome number evolution?
3) What are the patterns of petal color and petal number evolution, and are their 
evolution correlated?
Lastly, I discuss ways in which these analyses can be extended
to further examine the evolution of Onagraceae, 
and describe some recently developed metrics that can
be used to assess the informativeness of phylogenetic supermatrices.
